---
title: "Actinopterygii polymorhism analyses"
author: "Alina G. Mikhailova"
date: "2022-12-13"
output:    
   html_document:
      toc: true
      toc_depth: 3
      toc_float: true
      number_section: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, message=FALSE}
rm(list=ls(all=TRUE))
if (!require(caper)) install.packages("caper")
if (!require(geiger)) install.packages("geiger")
if (!require(ggpubr)) install.packages("ggpubr")
if (!require(dplyr)) install.packages("dplyr")
library(caper)
library(geiger)
library(ggpubr)
library(dplyr)
```

# Reading MutSpec DataBase with temperature and maturation time of Actinopterigii 

```{r}
mutSpec = read.table('../Data/3results/VertebratePolymorphisms.MutSpecData.txt', header = TRUE)
```

# Creating different subsets of mutspecs
## Subset of only CytB
```{r message=FALSE}
mutSpecCytB = mutSpec[mutSpec$Gene == "CytB",]
table(mutSpecCytB[!is.na(mutSpecCytB$temperature),]$Class)
```
## Subset of mutspec counted by mean normalised mutspec
```{r message=FALSE}
mutSpecAllMean = mutSpec %>% group_by(Species, Class, temperature, matur_tm); mutSpecAllMean = mutSpecAllMean %>% summarise(A_C.SubstRate.heavy=mean(A_C.SubstRate.heavy), A_G.SubstRate.heavy=mean(A_G.SubstRate.heavy), A_T.SubstRate.heavy=mean(A_T.SubstRate.heavy), C_A.SubstRate.heavy=mean(C_A.SubstRate.heavy), C_G.SubstRate.heavy=mean(C_G.SubstRate.heavy), C_T.SubstRate.heavy=mean(C_T.SubstRate.heavy), G_A.SubstRate.heavy=mean(G_A.SubstRate.heavy), G_C.SubstRate.heavy=mean(G_C.SubstRate.heavy), G_T.SubstRate.heavy=mean(G_T.SubstRate.heavy), T_A.SubstRate.heavy=mean(T_A.SubstRate.heavy), T_C.SubstRate.heavy=mean(T_C.SubstRate.heavy), T_G.SubstRate.heavy=mean(T_G.SubstRate.heavy))
table(mutSpecAllMean[!is.na(mutSpecAllMean$temperature),]$Class)
```


# Average MutSpec for Actinopteri
## CytB
```{r message=FALSE}
averageMutSpec1 = mutSpecCytB[!is.na(mutSpecCytB$temperature),][,26:37]
averageMutSpec1 = as.data.frame(apply(averageMutSpec1, 2, mean)); averageMutSpec1$Subs = c("A_G.h", "A_T.h", "A_C.h", "T_G.h", "T_A.h", "T_C.h", "G_T.h", "G_C.h", "G_A.h", "C_G.h", "C_T.h", "C_A.h"); names(averageMutSpec1) = c("Freq", "Subs")
f1 = ggbarplot(averageMutSpec1, x = "Subs", y = "Freq", fill = "Subs", color = "Subs",
          xlab="Substitution types", ylab="Normalised frequencies", legend = "none")  
f1
```

## Counted by mean normalised mutspec
```{r message=FALSE}
averageMutSpec2 = mutSpecAllMean[!is.na(mutSpecAllMean$temperature),][,5:16]
averageMutSpec2 = averageMutSpec2[!is.na(averageMutSpec2$C_A.SubstRate.heavy) & !is.na(averageMutSpec2$C_G.SubstRate.heavy) & !is.na(averageMutSpec2$C_T.SubstRate.heavy),]
averageMutSpec2 = as.data.frame(apply(averageMutSpec2, 2, mean)); averageMutSpec2$Subs = c("A_C.h", "A_G.h", "A_T.h", "C_A.h", "C_G.h",
 "C_T.h", "G_A.h", "G_C.h", "G_T.h", "T_A.h","T_C.h", "T_G.h"); names(averageMutSpec2) = c("Freq", "Subs")
pdf('../Figures/PolymorphicData.Actinopterygii.AverageMutSpec.pdf')
f1 = ggbarplot(averageMutSpec2, x = "Subs", y = "Freq", fill = "Subs", color = "Subs", palette = c("#bdbdbd", "#73514f", "#bdbdbd", "#bdbdbd", "#bdbdbd", "#055088", "#9c3d37", "#bdbdbd", "#bdbdbd", "#bdbdbd", "#036a5b", "#bdbdbd"),
          xlab="Substitution types", ylab="Normalised frequencies", legend = "none")  
f1
dev.off()
f1
```

# Correlations of MutSpec with temperature
## CytB

```{r message=FALSE, warning=FALSE} 
cor.test(mutSpecCytB$A_G.SubstRate.heavy,mutSpecCytB$temperature, method = 'spearman')   
cor.test(mutSpecCytB$G_A.SubstRate.heavy,mutSpecCytB$temperature, method = 'spearman')
cor.test(mutSpecCytB$T_C.SubstRate.heavy,mutSpecCytB$temperature, method = 'spearman')    
cor.test(mutSpecCytB$C_T.SubstRate.heavy,mutSpecCytB$temperature, method = 'spearman')
```

## Counted by mean normalised mutspec
```{r message=FALSE, warning=FALSE}  
mutSpecAllMean = mutSpecAllMean[!is.na(mutSpecAllMean$C_T.SubstRate.heavy),]
cor.test(mutSpecAllMean$A_G.SubstRate.heavy,mutSpecAllMean$temperature, method = 'spearman')   
cor.test(mutSpecAllMean$G_A.SubstRate.heavy,mutSpecAllMean$temperature, method = 'spearman')
cor.test(mutSpecAllMean$T_C.SubstRate.heavy,mutSpecAllMean$temperature, method = 'spearman')    
cor.test(mutSpecAllMean$C_T.SubstRate.heavy,mutSpecAllMean$temperature, method = 'spearman')   
```

# Correlation of AGdivTC and CTdivGA with temperature
## CYT B
```{r message=FALSE, warning=FALSE}
mutSpecCytB$AGdivTC=mutSpecCytB$A_G.SubstRate.heavy/mutSpecCytB$T_C.SubstRate.heavy
mutSpecCytB$CTdivGA=mutSpecCytB$C_T.SubstRate.heavy/mutSpecCytB$G_A.SubstRate.heavy
mutSpecCytB[mutSpecCytB$AGdivTC == "Inf" | mutSpecCytB$AGdivTC == "NaN",]$AGdivTC = NA
mutSpecCytB[mutSpecCytB$CTdivGA == "Inf" | mutSpecCytB$CTdivGA == "NaN",]$CTdivGA = NA
cor.test(mutSpecCytB$AGdivTC,mutSpecCytB$temperature, method = 'spearman')
cor.test(mutSpecCytB$CTdivGA,mutSpecCytB$temperature, method = 'spearman')
```
## MEAN

```{r message=FALSE, warning=FALSE}
mutSpecAllMean$AGdivTC=mutSpecAllMean$A_G.SubstRate.heavy/mutSpecAllMean$T_C.SubstRate.heavy
mutSpecAllMean$CTdivGA=mutSpecAllMean$C_T.SubstRate.heavy/mutSpecAllMean$G_A.SubstRate.heavy
mutSpecAllMean[mutSpecAllMean$AGdivTC == "Inf" | mutSpecAllMean$AGdivTC == "NaN",]$AGdivTC = NA
mutSpecAllMean[mutSpecAllMean$CTdivGA == "Inf" | mutSpecAllMean$CTdivGA == "NaN",]$CTdivGA = NA
cor.test(mutSpecAllMean$AGdivTC,mutSpecAllMean$temperature, method = 'spearman')
cor.test(mutSpecAllMean$CTdivGA,mutSpecAllMean$temperature, method = 'spearman')
```

## Figures
### Figures for mutspec all genes mean

```{r message=FALSE, warning= FALSE}
mutSpecAllMean = mutSpecAllMean[!is.na(mutSpecAllMean$temperature),]; N = as.character(paste("N", nrow(mutSpecAllMean), sep="")) #delete NA
f1a = ggscatter(mutSpecAllMean, x = "temperature", y = "A_G.SubstRate.heavy",
          color = "#73514f", # Points color, shape and size
          add = "reg.line",  # Add regressin line
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          conf.int = TRUE, # Add confidence interval
          yscale = "log2", xlab="Median annual water temperature, ?C", ylab="log2 AH>GH") + stat_cor(method = "spearman", aes(label = paste(..r.label.., ..p.label.., ..N.., sep = "~`,`~")))
f1a
```

```{r message=FALSE, warning= FALSE}
f1b1 = ggscatter(mutSpecAllMean, x = "temperature", y = "C_T.SubstRate.heavy",
          color = "#055088", # Points color, shape and size
          add = "reg.line",  # Add regressin line
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          conf.int = TRUE, # Add confidence interval
          yscale = "log2", xlab="Median annual water temperature, ?C", ylab="log2 CH>TH")+ stat_cor(method = "spearman", aes(label = paste(..r.label.., ..p.label.., ..N.., sep = "~`,`~")))
f1b1
```   

```{r message=FALSE, warning= FALSE}
f1b2 = ggscatter(mutSpecAllMean, x = "temperature", y = "G_A.SubstRate.heavy",
          color = "#9c3d37", # Points color, shape and size
          add = "reg.line",  # Add regressin line
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          conf.int = TRUE, # Add confidence interval
          yscale = "log2", xlab="Median annual water temperature, ?C", ylab="log2 GH>AH")+ stat_cor(method = "spearman", aes(label = paste(..r.label.., ..p.label.., ..N.., sep = "~`,`~")))
f1b2
```       

```{r message=FALSE, warning= FALSE}
f1c = ggscatter(mutSpecAllMean, x = "temperature", y = "AGdivTC",
          color = "#814194", # Points color, shape and size
          add = "reg.line",  # Add regressin line
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          conf.int = TRUE, # Add confidence interval
          yscale = "log2", xlab="Median annual water temperature, ?C", ylab="log2 A_GdivT_C")+ stat_cor(method = "spearman", aes(label = paste(..r.label.., ..p.label.., ..N.., sep = "~`,`~")))
f1c
``` 


# correlation of mutspec with time of maturation in fishes

```{r message=FALSE, warning=FALSE}  
cor.test(mutSpecAllMean$A_G.SubstRate.heavy,mutSpecAllMean$matur_tm, method = 'spearman')      
cor.test(mutSpecAllMean$G_A.SubstRate.heavy,mutSpecAllMean$matur_tm, method = 'spearman')    
cor.test(mutSpecAllMean$T_C.SubstRate.heavy,mutSpecAllMean$matur_tm, method = 'spearman')    
cor.test(mutSpecAllMean$C_T.SubstRate.heavy,mutSpecAllMean$matur_tm, method = 'spearman') 
```

# LM

## Cyt B
```{r}
summary(lm(formula = temperature ~ scale(A_G.SubstRate.heavy) + scale(C_T.SubstRate.heavy), data = mutSpecCytB))
summary(lm(formula = A_G.SubstRate.heavy ~ scale(temperature) * scale(matur_tm), data = mutSpecCytB))
summary(lm(formula = A_G.SubstRate.heavy ~ scale(temperature) + scale(matur_tm), data = mutSpecCytB))
summary(lm(formula = A_G.SubstRate.heavy ~ scale(temperature), data = mutSpecCytB))

mutSpecCytBNAzeroomit = mutSpecCytB[!is.na(mutSpecCytB$AGdivTC),]
mutSpecCytBNAzeroomit = mutSpecCytBNAzeroomit[mutSpecCytBNAzeroomit$AGdivTC != 0,]

summary(lm(formula = AGdivTC ~ scale(temperature) * scale(matur_tm), data = mutSpecCytBNAzeroomit))
summary(lm(formula = AGdivTC ~ scale(temperature) + scale(matur_tm), data = mutSpecCytBNAzeroomit))
summary(lm(formula = AGdivTC ~ scale(temperature), data = mutSpecCytBNAzeroomit))
``` 
## Mean 
```{r}
summary(lm(formula = temperature ~ scale(A_G.SubstRate.heavy) + scale(C_T.SubstRate.heavy), data = mutSpecAllMean))
summary(lm(formula = matur_tm ~ scale(A_G.SubstRate.heavy) + scale(C_T.SubstRate.heavy), data = mutSpecAllMean))
summary(lm(formula = A_G.SubstRate.heavy ~ scale(temperature) * scale(matur_tm), data = mutSpecAllMean))
summary(lm(formula = A_G.SubstRate.heavy ~ scale(temperature) + scale(matur_tm), data = mutSpecAllMean))
summary(lm(formula = A_G.SubstRate.heavy ~ scale(temperature), data = mutSpecAllMean))

mutSpecAllMeanNAzeroomit = mutSpecAllMean[!is.na(mutSpecAllMean$AGdivTC),]
mutSpecAllMeanNAzeroomit = mutSpecAllMeanNAzeroomit[mutSpecAllMeanNAzeroomit$AGdivTC != 0,]

summary(lm(formula = AGdivTC ~ scale(temperature) * scale(matur_tm), data = mutSpecAllMeanNAzeroomit))
summary(lm(formula = AGdivTC ~ scale(temperature) + scale(matur_tm), data = mutSpecAllMeanNAzeroomit))
summary(lm(formula = AGdivTC ~ scale(temperature), data = mutSpecAllMeanNAzeroomit))

summary(lm(formula = C_T.SubstRate.heavy ~ scale(temperature) * scale(matur_tm), data = mutSpecAllMean))
summary(lm(formula = C_T.SubstRate.heavy ~ scale(temperature) + scale(matur_tm), data = mutSpecAllMean))
summary(lm(formula = C_T.SubstRate.heavy ~ scale(temperature), data = mutSpecAllMean))

``` 

# PGLS 

## CytB
```{r message=FALSE, warning=FALSE}
tree = read.tree('../Data/1raw/mtalign.aln.treefile.rooted')

row.names(mutSpecCytBNAzeroomit) = mutSpecCytBNAzeroomit$Species
tree_pruned = treedata(tree, mutSpecCytBNAzeroomit, sort=T, warnings=T)$phy

data<-as.data.frame(treedata(tree_pruned, mutSpecCytBNAzeroomit, sort=T, warnings=T)$data)
data$Species = as.character(data$Species)
data$AGdivTC = as.numeric(as.character(data$AGdivTC))
data$temperature = as.numeric(as.character(data$temperature))
data$matur_tm = as.numeric(as.character(data$matur_tm))
data_comp <- comparative.data(tree_pruned, data, Species, vcv=TRUE)

summary(pgls(AGdivTC ~ scale(temperature), data_comp, lambda="ML"))
summary(pgls(log2(AGdivTC) ~ log2(temperature), data_comp, lambda="ML"))
```

## Mean
```{r message=FALSE, warning=FALSE}
tree = read.tree('../Data/1raw/mtalign.aln.treefile.rooted')

row.names(mutSpecAllMeanNAzeroomit) = mutSpecAllMeanNAzeroomit$Species
tree_pruned = treedata(tree, mutSpecAllMeanNAzeroomit, sort=T, warnings=T)$phy

data<-as.data.frame(treedata(tree_pruned, mutSpecAllMeanNAzeroomit, sort=T, warnings=T)$data)
data$Species = as.character(data$Species)
data$AGdivTC = as.numeric(as.character(data$AGdivTC))
data$temperature = as.numeric(as.character(data$temperature))
data$matur_tm = as.numeric(as.character(data$matur_tm))
data_comp <- comparative.data(tree_pruned, data, Species, vcv=TRUE)

summary(pgls(AGdivTC ~ scale(temperature), data_comp, lambda="ML"))
summary(pgls(log2(AGdivTC) ~ log2(temperature), data_comp, lambda="ML"))
```