---
title: "AnalysesGraphsForSupplements"
author: "Dmitrii Iliushchenko"
date: "2022-12-14"
output:
      html_document:
      toc: true
      toc_depth: 3
      toc_float: true
      number_section: true
---

```{r setup, echo=TRUE, message=FALSE}
rm(list=ls(all=TRUE))

library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
library(ggplot2)
```

### Read MutSpec with temperature and taxa
```{r}
muttemp = read.table('../data/3results/VertebratePolymorphisms.MutSpecData.txt', header=T)

```

### Add families from taxa df
```{r}
taxa = read.table("../data/1raw/TaxaFromKostya.Names.stat", sep = '\t',header = FALSE) 
taxa = as.data.frame(taxa[grepl('Actinopteri',taxa$V1),])
names(taxa) = 'taxa'

taxa$Species = gsub(";(.*)",'',taxa$taxa);
taxa$Species = gsub(" ",'_',taxa$Species);
taxa$Family = gsub(";Actinopteri(.*)",'',taxa$taxa)
taxa$Family = gsub("(.*);",'',taxa$Family)
taxa = taxa[,-1]

Fishes = merge(muttemp, taxa, by = 'Species')
for (i in 1:nrow(Fishes))   {Fishes$FamilyShort[i] = paste(unlist(strsplit(Fishes$Family[i],''))[c(1:3)],collapse = '') }

```

### Take only fishes by temperature 
```{r}
Fishes = Fishes[!is.na(Fishes$temperature),]
```

### Calculate TsTv metric
```{r}
Fishes$TsTv = (Fishes$T_C.SubstRate.heavy + Fishes$C_T.SubstRate.heavy + Fishes$G_A.SubstRate.heavy + Fishes$A_G.SubstRate.heavy) / 
              (Fishes$T_A.SubstRate.heavy + Fishes$A_T.SubstRate.heavy + Fishes$G_C.SubstRate.heavy + Fishes$C_G.SubstRate.heavy + 
               Fishes$G_T.SubstRate.heavy + Fishes$T_G.SubstRate.heavy + Fishes$C_A.SubstRate.heavy + Fishes$A_C.SubstRate.heavy)
```

###Ddelete families, that have less than 3 species!!!
```{r}
FamFreq = data.frame(table(Fishes$Family))
FrequentFamilies = FamFreq[FamFreq$Freq >= 3,]$Var1
FamFreq = FamFreq[FamFreq$Var1 %in% FrequentFamilies,]
names(FamFreq)=c('Family','NumberOfSpecies')
Fishes = Fishes[Fishes$Family %in% FrequentFamilies,]
```

###  Calculate Ts/Tv and A<G/T<C
```{r}
Fishes$A_G.T_C.heavy = (Fishes$A_G.heavy/Fishes$T_C.heavy)
Fishes = Fishes[Fishes$A_G.T_C.heavy > 0,]
Fishes = Fishes[Fishes$A_G.T_C.heavy < Inf,]
Fishes = Fishes[Fishes$TsTv < Inf,]
Fishes = Fishes[!is.na(Fishes$A_G.T_C.heavy),]
```

### Calculate median of A>G/T>C; TsTv; Temperature by Families
```{r}
agg = aggregate(list(Fishes$TsTv, Fishes$A_G.T_C.heavy, Fishes$temperature), by = list(Fishes$Family), FUN = median)
names(agg) = c('Family','TsTv','A_G.T_C','Temperature')
for (i in 1:nrow(agg))   {agg$FamilyShort[i] = paste(unlist(strsplit(agg$Family[i],''))[c(1:3)],collapse = '')  }
```

### Cor test A>G/T>C on heavy chain with temperature
```{r}
cor.test(agg$A_G.T_C, agg$Temperature,method = 'spearman')
```
### Plots
```{r}
ggplot(data = agg, aes(x = Temperature, y = A_G.T_C, label = FamilyShort, color = Family))+
  geom_point()+
  geom_smooth(method="lm", se=F, col = 'red')+
  theme_classic()+
  labs(x = 'Median annual water temperature, °C', y = 'Median A>G/T>C')+
  geom_text(aes(label=FamilyShort),hjust=-0.15, vjust=-0.5, show.legend = F)
  
```

### Plots for Families
```{r}
ggplot(data = Fishes, aes(x = temperature, y = A_G.T_C.heavy, group = FamilyShort, fill = FamilyShort))+
  geom_boxplot(alpha=0.3)+
  theme_classic()+
  labs(x = 'Temperature, C°', y = 'A>G/T>C')
```

### Boxplots by quartiles
```{r}
boxplot(Fishes[Fishes$temperature<=quantile(Fishes$temperature,0.25),]$A_G.T_C,
        Fishes[Fishes$temperature>quantile(Fishes$temperature,0.25) & Fishes$temperature<=quantile(Fishes$temperature,0.5),]$A_G.T_C,
        Fishes[Fishes$temperature>quantile(Fishes$temperature,0.5) & Fishes$temperature<=quantile(Fishes$temperature,0.75),]$A_G.T_C,
        Fishes[Fishes$temperature>quantile(Fishes$temperature,0.75),]$A_G.T_C,
        names=c('1','2','3','4'), outline = FALSE, notch = FALSE, ylab = 'A>G/T>C', xlab = 'Quartiles of median annual water temperature, C')

```

### Boxplots by median

```{r}
boxplot(Fishes[Fishes$temperature<=quantile(Fishes$temperature,0.5),]$A_G.T_C,
        Fishes[Fishes$temperature>quantile(Fishes$temperature,0.5),]$A_G.T_C,
        names=c('colder fishes','warmer fishes'), outline = FALSE, notch = TRUE, cex = 3, ylab = 'A>G/T>C')
```

### Define fishes by temp and maturation
```{r}
fishes_matur = muttemp[!is.na(muttemp$matur_tm),]
fishes_matur = fishes_matur[!is.na(fishes_matur$temperature),]
```

### Calculate metrics 
```{r}
fishes_matur$CtoTSkew = (fishes_matur$FrC.heavy-fishes_matur$FrT.heavy)/(fishes_matur$FrC.heavy+fishes_matur$FrT.heavy) 
fishes_matur$GtoASkew = (fishes_matur$FrG.heavy-fishes_matur$FrA.heavy)/(fishes_matur$FrG.heavy+fishes_matur$FrA.heavy)
fishes_matur$AtoGSkew = (fishes_matur$FrA.heavy-fishes_matur$FrG.heavy)/(fishes_matur$FrA.heavy+fishes_matur$FrG.heavy)
fishes_matur$TtoCSkew = (fishes_matur$FrT.heavy-fishes_matur$FrC.heavy)/(fishes_matur$FrT.heavy+fishes_matur$FrC.heavy)
fishes_matur$TG = fishes_matur$FrT.heavy+fishes_matur$FrG.heavy
fishes_matur$AC = fishes_matur$FrA.heavy+fishes_matur$FrC.heavy
fishes_matur$TG_ACSkew = (fishes_matur$TG-fishes_matur$AC)/(fishes_matur$TG+fishes_matur$AC)
fishes_matur$AC_TGSkew = -(fishes_matur$TG-fishes_matur$AC)/(fishes_matur$TG+fishes_matur$AC)
```

### Function to define long or short maturated
```{r}
medmatur = median(fishes_matur$matur_tm)

mmm = function(x)
{
  if (x < medmatur) {return('Short Maturated')}
  else{return('Long Maturated')}
}
```

### Define long or short maturated
```{r}
fishes_matur$Maturated = apply(as.matrix(fishes_matur$matur_tm),1,mmm)

m_temp_s = median(fishes_matur[fishes_matur$Maturated == 'Short Maturated',]$temperature)
m_temp_l = median(fishes_matur[fishes_matur$Maturated == 'Long Maturated',]$temperature)
```


### Define what cold or warm incide short and long maturated fishes
```{r}
for (i in 1:nrow(fishes_matur))
{
  if (fishes_matur$Maturated[i] == 'Short Maturated')
  {
    if (fishes_matur$temperature[i] <= m_temp_s) {fishes_matur$Temp[i] = 'Colder Fishes'}
    else {fishes_matur$Temp[i] = 'Warmer Fishes'}
  }
  else if (fishes_matur$Maturated[i] == 'Long Maturated') 
  {
    if (fishes_matur$temperature[i] <= m_temp_l) {fishes_matur$Temp[i] = 'Colder Fishes'}
    else {fishes_matur$Temp[i] = 'Warmer Fishes'}
  }
}
```


### Boxplots of maturated and cold/warm fishes
```{r}
ggplot(data = fishes_matur, aes(x = Temp, y = TG_ACSkew))+
  geom_boxplot()+
  facet_wrap(~Maturated)+
  theme_test()+
  labs(y = 'TGh-ACh skew')
```



