---
title: "CriticalThinkingAgainKp.PleaseDeleteMeLater.Rmd"
author: "Konstantin Popadin"
date: "11 Dec 2022"
output: html_document
---

```{r setup, include=FALSE}
library(caper)
library(geiger)
library(ggplot2)
library(dplyr)
```
```{r echo=FALSE}
rm(list=ls(all=TRUE))
```
### 1: Reading MutSpec and extraction of Cold and Warm MutSpecs of Actinopterygii for CytB with known temperature. How I would start simulations
```{r}

mutSpec = read.table('../Data/3results/VertebratePolymorphisms.MutSpecData.txt', header = TRUE)
mutSpec = mutSpec[mutSpec$Class == 'Actinopterygii',]
mutSpec = mutSpec[mutSpec$Gene == 'CytB',]
summary(mutSpec$temperature)
mutSpec = mutSpec[!is.na(mutSpec$temperature),]
nrow(mutSpec)

```
### 2: Substitution rates positively correlate with the number of sequences. It is not good. Number of Sequences should not.
```{r}
cor.test(mutSpec$T_C.N,mutSpec$NumOfFourFoldMut, method = 'spearman') # Ah>Gh
cor.test(mutSpec$T_C.N,mutSpec$NumberOfSeq, method = 'spearman')      # Ah>Gh

A<-lm(mutSpec$T_C.N ~ scale(mutSpec$temperature) + scale(mutSpec$NumOfFourFoldMut))
summary(A)
B<-lm(mutSpec$T_C.N ~ scale(mutSpec$temperature) + scale(mutSpec$NumberOfSeq))
summary(B)

```
### 3: Rederive the MutSpec: 
### => Different normalisation: divide by the number of neutral nucletodes in the concatenate of all external sequences;
### => Heavy chain notation.
```{r}
mutSpec$A_G.SubstRate.heavy = mutSpec$T_C/mutSpec$T
mutSpec$A_T.SubstRate.heavy = mutSpec$T_A/mutSpec$T
mutSpec$A_C.SubstRate.heavy = mutSpec$T_G/mutSpec$T
mutSpec$T_G.SubstRate.heavy = mutSpec$A_C/mutSpec$A
mutSpec$T_A.SubstRate.heavy = mutSpec$A_T/mutSpec$A
mutSpec$T_C.SubstRate.heavy = mutSpec$A_G/mutSpec$A
mutSpec$G_T.SubstRate.heavy = mutSpec$C_A/mutSpec$C
mutSpec$G_C.SubstRate.heavy = mutSpec$C_G/mutSpec$C
mutSpec$G_A.SubstRate.heavy = mutSpec$C_T/mutSpec$C
mutSpec$C_G.SubstRate.heavy = mutSpec$G_C/mutSpec$G
mutSpec$C_T.SubstRate.heavy = mutSpec$G_A/mutSpec$G
mutSpec$C_A.SubstRate.heavy = mutSpec$G_T/mutSpec$G
mutSpec$TotalSubstRate.heavy = mutSpec$A_G.SubstRate.heavy + mutSpec$A_T.SubstRate.heavy + mutSpec$A_C.SubstRate.heavy
+ mutSpec$T_G.SubstRate.heavy + mutSpec$T_A.SubstRate.heavy + mutSpec$T_C.SubstRate.heavy
+ mutSpec$G_T.SubstRate.heavy + mutSpec$G_C.SubstRate.heavy + mutSpec$G_A.SubstRate.heavy
+ mutSpec$C_G.SubstRate.heavy + mutSpec$C_T.SubstRate.heavy + mutSpec$C_A.SubstRate.heavy

mutSpec$TotalTs.SubstRate.heavy = mutSpec$A_G.SubstRate.heavy + mutSpec$G_A.SubstRate.heavy + mutSpec$C_T.SubstRate.heavy +  mutSpec$T_C.SubstRate.heavy

mutSpec$TotalTv.SubstRate.heavy = mutSpec$A_T.SubstRate.heavy + mutSpec$A_C.SubstRate.heavy + mutSpec$T_G.SubstRate.heavy + mutSpec$T_A.SubstRate.heavy + mutSpec$G_T.SubstRate.heavy + mutSpec$G_C.SubstRate.heavy + mutSpec$C_G.SubstRate.heavy + mutSpec$C_A.SubstRate.heavy

```
### 4: Check of the integral MutSpec, 
### Substitution rates do not correlate anymore with the number of sequences (it is good - more correct normalization);
### Substitution rates correlate with the number of 4F mutations. Expectedly!?
### Substitution rates correlate with temperature (even the total substitution rate) - higher global mut rate and within this trend A>G is the strongest!?
```{r}
## integral MutSpec looks good:
barplot(c(mean(mutSpec$A_G.SubstRate.heavy),mean(mutSpec$A_T.SubstRate.heavy),mean(mutSpec$A_C.SubstRate.heavy),
mean(mutSpec$T_G.SubstRate.heavy),mean(mutSpec$T_A.SubstRate.heavy),mean(mutSpec$T_C.SubstRate.heavy),
mean(mutSpec$G_T.SubstRate.heavy),mean(mutSpec$G_C.SubstRate.heavy),mean(mutSpec$G_A.SubstRate.heavy),
mean(mutSpec$C_G.SubstRate.heavy),mean(mutSpec$C_T.SubstRate.heavy),mean(mutSpec$C_A.SubstRate.heavy)), names = c('Ah>Gh','Ah>Th','Ah>Ch','Th>Gh','Th>Ah','Th>Ch','Gh>Th','Gh>Ch','Gh>Ah','Ch>Gh','Ch>Th','Ch>Ah'), cex.names = 0.65)

## if there is a correlation of substitution rates with the number of sequences? No, because normalization is OK now
cor.test(mutSpec$TotalSubstRate.heavy, mutSpec$NumberOfSeq, method = 'spearman') # no
cor.test(mutSpec$A_G.SubstRate.heavy, mutSpec$NumberOfSeq, method = 'spearman') # no
cor.test(mutSpec$C_T.SubstRate.heavy, mutSpec$NumberOfSeq, method = 'spearman') # no

## if there is a correlation between the  substitution rates and NumberOfFourFoldMut? Yes, sure
cor.test(mutSpec$TotalSubstRate.heavy, mutSpec$NumOfFourFoldMut, method = 'spearman') # very positive
cor.test(mutSpec$A_G.SubstRate.heavy, mutSpec$NumOfFourFoldMut, method = 'spearman') # very positive
cor.test(mutSpec$C_T.SubstRate.heavy, mutSpec$NumOfFourFoldMut, method = 'spearman') # very positive

## total substitution rate - if there is correlation with temperature? yes. Warm water mutate faster in general and Ah>Gh is the best among them!?
cor.test(mutSpec$TotalSubstRate.heavy, mutSpec$temperature, method = 'spearman') # a bit positive
cor.test(mutSpec$A_G.SubstRate.heavy, mutSpec$temperature, method = 'spearman') # positive
cor.test(mutSpec$C_T.SubstRate.heavy, mutSpec$temperature, method = 'spearman') # a bit positive
cor.test(mutSpec$T_C.SubstRate.heavy, mutSpec$temperature, method = 'spearman') # nothing
cor.test(mutSpec$G_A.SubstRate.heavy, mutSpec$temperature, method = 'spearman') # a bit positive

cor.test(mutSpec$TotalTs.SubstRate.heavy, mutSpec$temperature, method = 'spearman') # a bit positive
cor.test(mutSpec$TotalTv.SubstRate.heavy, mutSpec$temperature, method = 'spearman') # about to be absent

### A>G as a function of temperature and total number of substitutions
A<-lm(mutSpec$A_G.SubstRate.heavy ~ mutSpec$temperature + mutSpec$TotalSubstRate.heavy); summary(A)
A<-lm(mutSpec$A_G.SubstRate.heavy ~ mutSpec$temperature); summary(A)
A<-lm(mutSpec$A_G.SubstRate.heavy/mutSpec$TotalSubstRate.heavy ~ mutSpec$temperature); summary(A)

B<-lm(mutSpec$C_T.SubstRate.heavy ~ mutSpec$temperature + mutSpec$TotalSubstRate.heavy); summary(B)
B<-lm(mutSpec$C_T.SubstRate.heavy ~ mutSpec$temperature); summary(B)
# B<-lm(mutSpec$C_T.SubstRate.heavy/mutSpec$TotalSubstRate.heavy ~ mutSpec$temperature); summary(B) # Inf sometimes...

### backward multiple model
C<-lm(mutSpec$temperature ~ mutSpec$A_G.SubstRate.heavy + mutSpec$C_T.SubstRate.heavy + mutSpec$TotalSubstRate.heavy + mutSpec$TotalTs.SubstRate.heavy + mutSpec$TotalTv.SubstRate.heavy); summary(C)
C<-lm(mutSpec$temperature ~ mutSpec$A_G.SubstRate.heavy + mutSpec$C_T.SubstRate.heavy + mutSpec$TotalTs.SubstRate.heavy + mutSpec$TotalTv.SubstRate.heavy); summary(C)
C<-lm(mutSpec$temperature ~ mutSpec$A_G.SubstRate.heavy + mutSpec$C_T.SubstRate.heavy + mutSpec$TotalTs.SubstRate.heavy); summary(C)
C<-lm(mutSpec$temperature ~ mutSpec$A_G.SubstRate.heavy + mutSpec$TotalTs.SubstRate.heavy); summary(C)
C<-lm(mutSpec$temperature ~ mutSpec$A_G.SubstRate.heavy); summary(C)
C<-lm(mutSpec$temperature ~ mutSpec$A_G.SubstRate.heavy/mutSpec$T_C.SubstRate.heavy); summary(C) ### DIGEST IT!!!!!!!!!!

### prcomp!!! TO DO!


```

