---
title: "Derive Mutational spectrum"
author: "Dmitrii Iliushchenko"
date: "15/9/2022"
output:
    html_document:
      toc: true
      toc_depth: 3
      toc_float: true
      number_section: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

```{r echo=FALSE, message=FALSE}
rm(list=ls(all=TRUE))

library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
```
# Get Neutral ATGC for each species and gene

## Preparing 

### Define Four Fold sites
```{r}
VecOfSynFourFoldDegenerateSites <- c('CTT', 'CTC', 'CTA', 'CTG', 
                                     'GTT', 'GTC', 'GTA', 'GTG', 
                                     'TCT', 'TCC', 'TCA', 'TCG', 
                                     'CCT', 'CCC', 'CCA', 'CCG', 
                                     'ACT', 'ACC', 'ACA', 'ACG', 
                                     'GCT', 'GCC', 'GCA', 'GCG', 
                                     'CGT', 'CGC', 'CGA', 'CGG', 
                                     'GGT', 'GGC', 'GGA', 'GGG')
```

### Unzip (optionaly) files and get list of them
```{r}
#unzip("../Data/1raw/POLARIZEDBR_DATA.zip")
List = list.files("../Data/1raw/POLARIZEDBR_DATA/")
```

## Parse into one df and count neutral ATGC
```{r}
for (i in 1:length(List)){
  infile = paste("../Data/1raw/POLARIZEDBR_DATA/",as.character(List[i]),sep='') 
  if (length(grep('POLARISED',infile)) > 0)
  {
    Species = gsub('\\..*','',as.character(List[i]))
    Gene = gsub(Species,'',as.character(List[i])); Gene = gsub('.POLARISED.txt','',Gene); Gene = gsub('\\.POLARISED.txt','',Gene); Gene = gsub('\\.','',Gene); 
    GeneSpecies = read.table(infile, header = TRUE)
    GeneSpecies = GeneSpecies[GeneSpecies$BranchPosition == 'External',]
    NumOfSeqSpXG = nrow(GeneSpecies)
    ExternalSeqsTogether = paste(GeneSpecies$MoreShallowNodeSeq,collapse = '')
    ExternalSeqsTogether = unlist(strsplit(ExternalSeqsTogether,'')) # 5700/3
    CodonsVec = c(); StartNuc = 1
    if (length(ExternalSeqsTogether)/3 == round(length(ExternalSeqsTogether)/3))  # if divide by 3 without the rest
    {
      for (j in 1:(length(ExternalSeqsTogether)/3))
      {
        CodonsVec = c(CodonsVec,paste(ExternalSeqsTogether[StartNuc : (StartNuc+2)],collapse = ''))
        StartNuc = StartNuc+3
      }
      AllCodons = length(CodonsVec)        # 1021
      CodonsVecNeutral = CodonsVec[CodonsVec %in% VecOfSynFourFoldDegenerateSites]
      NeutralCodons = length(CodonsVecNeutral) # 1900
      data.frame(table(CodonsVecNeutral))
      
      CodonsVecNeutral = gsub("CTA|GTA|TCA|CCA|ACA|GCA|CGA|GGA",'A',CodonsVecNeutral)
      CodonsVecNeutral = gsub("CTT|GTT|TCT|CCT|ACT|GCT|CGT|GGT",'T',CodonsVecNeutral)
      CodonsVecNeutral = gsub("CTG|GTG|TCG|CCG|ACG|GCG|CGG|GGG",'G',CodonsVecNeutral)
      CodonsVecNeutral = gsub("CTC|GTC|TCC|CCC|ACC|GCC|CGC|GGC",'C',CodonsVecNeutral)
      
      Line=c(Species,Gene,length(CodonsVecNeutral[CodonsVecNeutral == 'A']),
             length(CodonsVecNeutral[CodonsVecNeutral == 'T']),
             length(CodonsVecNeutral[CodonsVecNeutral == 'G']),
             length(CodonsVecNeutral[CodonsVecNeutral == 'C']),
             AllCodons, NeutralCodons, NumOfSeqSpXG)
      if (i == 1) {Final = Line}
      if (i >  1) {Final = rbind(Final,Line)}
    }
  }
}
Final = as.data.frame(Final); names(Final)=c('Species','Gene','A','T','G','C','NumberOfAllCodons',"NumberOfFourFoldDegenCodons",'NumberOfSeq')
```


## Merge with Taxa

### Prepare Taxa before Merging

In both df we cut out third word of sps

```{r}

Taxa = read.table("../Data/1raw/TaxaFromKostya.Names.stat", sep = '\t',header = FALSE) 
Taxa$Species = gsub(";.*",'',Taxa$V1); 
for (i in (1:nrow(Taxa)))  {Taxa$Species[i] = paste(unlist(strsplit(as.character(Taxa$Species[i]),split = ' '))[1],unlist(strsplit(as.character(Taxa$Species[i]),split = ' '))[2], sep = '_')}
Taxa$Class = gsub(";Chordata;.*",'',Taxa$V1); Taxa$Class = gsub(".*;",'',Taxa$Class); table(Taxa$Class)
Taxa$Class = gsub('Actinopteri','Actinopterygii',Taxa$Class)
Taxa$Class = gsub("Testudines|Squamata|Crocodylia|Sphenodontia",'Reptilia',Taxa$Class)
length(unique(Taxa$Species)) # 1708
table(Taxa$Class)
Taxa = Taxa[,-1]

TaxaMore = read.table("../Data/1raw/TaxaFromKostya.2NeedTaxa.tax.txt", sep = '\t',header = FALSE) 
TaxaMore$Species = ''
for (i in (1:nrow(TaxaMore)))  
{TaxaMore$Species[i] = paste(unlist(strsplit(as.character(TaxaMore$V1[i]),split = ' '))[1],
                             unlist(strsplit(as.character(TaxaMore$V1[i]),split = ' '))[2], sep = '_')}
TaxaMore$Class = gsub("; Chordata;.*",'',TaxaMore$V2); 
TaxaMore$Class = gsub(".*; ",'',TaxaMore$Class); 
TaxaMore$Class = gsub('Actinopteri','Actinopterygii',TaxaMore$Class)
TaxaMore$Class = gsub("Testudines|Squamata|Crocodylia",'Reptilia',TaxaMore$Class)
table(TaxaMore$Class)
TaxaMore = TaxaMore[,-c(1,2)]

Taxa = rbind(Taxa,TaxaMore); Taxa = unique(Taxa)
```

### Merging subs with taxa, if no taxa detected still hold it with NA
```{r}
Final = merge(Final,Taxa, all = TRUE, by='Species')
NeutralATGC = Final
```

# Read another file with mutational spectra

Here we have **2404 unique species**
```{r}
MUT = read.table("../Data/1raw/Mutational_spectra_in_Chordata_ML.txt", header = TRUE)
length(unique(MUT$Species)) 
```

# Prepare MUT df to future merging

## Filtration

### Take only normal substitutions and filter out species with too high fraction (> 5%) of unnormal substitutions

Define normal subs
```{r}
VecOfNormalSubstitutions <- c('A_C','C_A',
                              'A_G','G_A',
                              'C_G','G_C',
                              'C_T','T_C',
                              'G_T','T_G',
                              'T_A','A_T')
```

Than, select correct one
```{r}
SP = data.frame(table(MUT$Species)); names(SP) = c('Species','NumberOfAllSubst')
SPN = data.frame(table(MUT[MUT$Subs %in% VecOfNormalSubstitutions,]$Species)); names(SPN) = c('Species','NumberOfNormalSubst')
SP = merge(SP,SPN); SP$FractionOfNormal = SP$NumberOfNormalSubst/SP$NumberOfAllSubst

summary(SP$FractionOfNormal) # how many to delete? ask to have more than 95% of substitutions as normal
SpeciesToDelete = SP[SP$FractionOfNormal <=0.95,]$Species; length(SpeciesToDelete) # 279 - delete
MUT = MUT[!MUT$Species %in% SpeciesToDelete,]
MUT = MUT[MUT$Subs %in% VecOfNormalSubstitutions,]
```

### Filter Only subs occured in 4fold

Here we have mutations before and after filter
```{r}
nrow(MUT)
MUT = MUT[MUT$AncestralAA == MUT$DescendantAA,]
MUT4fold = MUT[MUT$AncestorCodon %in% VecOfSynFourFoldDegenerateSites & MUT$DescendantCodon %in% VecOfSynFourFoldDegenerateSites,]; 
MUT <- NULL
MUT4fold$Number = 1
nrow(MUT4fold)
```

# Counting and normalization

## Count number of different mutations
```{r}
AGG = aggregate(MUT4fold$Number, by = list(MUT4fold$Species, MUT4fold$Gene), FUN = sum)
names(AGG) = c('Species','Gene','NumOfFourFoldMut')
AGG[is.na(AGG)] <- 0
MUT4fold = merge(MUT4fold,AGG, by = c('Species', "Gene"))
MUT4fold = merge(MUT4fold,NeutralATGC[,c(1:6)], by = c("Species", "Gene"))
NeutralATGC = NULL
```

## Calculate number of subs in each species and gene
```{r}
mut_count = aggregate(MUT4fold$Number, by=list(MUT4fold$Species, MUT4fold$Gene, MUT4fold$Subs), FUN=sum); 
names(mut_count) = c('Species','Gene','Subs','Freq')
mut_count =
  mut_count %>% 
  pivot_wider(names_from = Subs, values_from = Freq, values_fill = 0)

MUT4fold = merge(MUT4fold,mut_count, by = c('Species', "Gene"))
```

## Normalization 12 subs
Normalizing by divide substitution on number of all substitution occurred in each species and gene

```{r echo = FALSE}

MUT4fold$A = as.numeric(MUT4fold$A)
MUT4fold$T = as.numeric(MUT4fold$T)
MUT4fold$G = as.numeric(MUT4fold$G)
MUT4fold$C = as.numeric(MUT4fold$C)

```

### Calculating FrX for 4 nucl
```{r}
MUT4fold$FrA = MUT4fold$A / (MUT4fold$A + MUT4fold$T + MUT4fold$C + MUT4fold$G)
MUT4fold$FrT = MUT4fold$T / (MUT4fold$A + MUT4fold$T + MUT4fold$C + MUT4fold$G)
MUT4fold$FrG = MUT4fold$G / (MUT4fold$A + MUT4fold$T + MUT4fold$C + MUT4fold$G)
MUT4fold$FrC = MUT4fold$C / (MUT4fold$A + MUT4fold$T + MUT4fold$C + MUT4fold$G)
```

### Normalization
```{r}
MUT4fold$A_C.N = MUT4fold$A_C / (MUT4fold$FrA * 100)
MUT4fold$A_G.N = MUT4fold$A_G / (MUT4fold$FrA * 100)
MUT4fold$A_T.N = MUT4fold$A_T / (MUT4fold$FrA * 100)
MUT4fold$C_A.N = MUT4fold$C_A / (MUT4fold$FrC * 100)
MUT4fold$C_G.N = MUT4fold$C_G / (MUT4fold$FrC * 100)
MUT4fold$C_T.N = MUT4fold$C_T / (MUT4fold$FrC * 100)
MUT4fold$G_A.N = MUT4fold$G_A / (MUT4fold$FrG * 100)
MUT4fold$G_C.N = MUT4fold$G_C / (MUT4fold$FrG * 100)
MUT4fold$G_T.N = MUT4fold$G_T / (MUT4fold$FrG * 100)
MUT4fold$T_A.N = MUT4fold$T_A / (MUT4fold$FrT * 100)
MUT4fold$T_C.N = MUT4fold$T_C / (MUT4fold$FrT * 100)
MUT4fold$T_G.N = MUT4fold$T_G / (MUT4fold$FrT * 100)
```

### Drop unused colums
```{r}
MUT_final = subset(MUT4fold, select = -c(AncestralSeqName, DescendantSeqName,Branch, Number))
MUT_final %>% 
  head() %>% 
  kable() %>% 
  kable_styling(full_width = F)
```
### Save table
```{r}
write.table(MUT_final, file = '../Data/3results/VertebratePolymorphisms.MutSpecData.txt', quote = FALSE)
```
