---
title: "EquilibriumFishes"
author: "Dmitrii Iliushchenko"
date: "2022-11-25"
output: html_document
---

```{r setup, include=FALSE}
library(caper)
library(geiger)
library(ggplot2)
library(dplyr)
```
```{r echo=FALSE}
rm(list=ls(all=TRUE))
```

### Reading temperature for fishes and mutspec

```{r message=FALSE}
muttemp = read.table('../data/3results/VertebratePolymorphisms.MutSpecData.txt', header =TRUE) 

mutSpecAllMean = muttemp %>% group_by(Species, Class, temperature, matur_tm); 
mutSpecAllMean = mutSpecAllMean %>% 
  summarise(FrA.heavy=mean(FrA.heavy),
            FrT.heavy=mean(FrT.heavy),
            FrG.heavy=mean(FrG.heavy),
            FrC.heavy=mean(FrC.heavy),
            A_C.SubstRate.heavy=mean(A_C.SubstRate.heavy), 
            A_G.SubstRate.heavy=mean(A_G.SubstRate.heavy), 
            A_T.SubstRate.heavy=mean(A_T.SubstRate.heavy), 
            C_A.SubstRate.heavy=mean(C_A.SubstRate.heavy), 
            C_G.SubstRate.heavy=mean(C_G.SubstRate.heavy), 
            C_T.SubstRate.heavy=mean(C_T.SubstRate.heavy), 
            G_A.SubstRate.heavy=mean(G_A.SubstRate.heavy), 
            G_C.SubstRate.heavy=mean(G_C.SubstRate.heavy), 
            G_T.SubstRate.heavy=mean(G_T.SubstRate.heavy), 
            T_A.SubstRate.heavy=mean(T_A.SubstRate.heavy), 
            T_C.SubstRate.heavy=mean(T_C.SubstRate.heavy), 
            T_G.SubstRate.heavy=mean(T_G.SubstRate.heavy))

mutSpecAllMean$subsratesum = mutSpecAllMean$A_C.SubstRate.heavy + 
                                 mutSpecAllMean$A_G.SubstRate.heavy +
                                 mutSpecAllMean$A_T.SubstRate.heavy +
                                 mutSpecAllMean$C_A.SubstRate.heavy +
                                 mutSpecAllMean$C_G.SubstRate.heavy +
                                 mutSpecAllMean$C_T.SubstRate.heavy +
                                 mutSpecAllMean$G_A.SubstRate.heavy +
                                 mutSpecAllMean$G_C.SubstRate.heavy +
                                 mutSpecAllMean$G_T.SubstRate.heavy +
                                 mutSpecAllMean$T_A.SubstRate.heavy +
                                 mutSpecAllMean$T_C.SubstRate.heavy +
                                 mutSpecAllMean$T_G.SubstRate.heavy

mutSpecAllMean$A_C.SubstRate.Percent.heavy=mutSpecAllMean$A_C.SubstRate.heavy/mutSpecAllMean$subsratesum
mutSpecAllMean$A_G.SubstRate.Percent.heavy=mutSpecAllMean$A_G.SubstRate.heavy/mutSpecAllMean$subsratesum
mutSpecAllMean$A_T.SubstRate.Percent.heavy=mutSpecAllMean$A_T.SubstRate.heavy/mutSpecAllMean$subsratesum
mutSpecAllMean$C_A.SubstRate.Percent.heavy=mutSpecAllMean$C_A.SubstRate.heavy/mutSpecAllMean$subsratesum
mutSpecAllMean$C_G.SubstRate.Percent.heavy=mutSpecAllMean$C_G.SubstRate.heavy/mutSpecAllMean$subsratesum
mutSpecAllMean$C_T.SubstRate.Percent.heavy=mutSpecAllMean$C_T.SubstRate.heavy/mutSpecAllMean$subsratesum
mutSpecAllMean$G_A.SubstRate.Percent.heavy=mutSpecAllMean$G_A.SubstRate.heavy/mutSpecAllMean$subsratesum
mutSpecAllMean$G_C.SubstRate.Percent.heavy=mutSpecAllMean$G_C.SubstRate.heavy/mutSpecAllMean$subsratesum
mutSpecAllMean$G_T.SubstRate.Percent.heavy=mutSpecAllMean$G_T.SubstRate.heavy/mutSpecAllMean$subsratesum
mutSpecAllMean$T_A.SubstRate.Percent.heavy=mutSpecAllMean$T_A.SubstRate.heavy/mutSpecAllMean$subsratesum
mutSpecAllMean$T_C.SubstRate.Percent.heavy=mutSpecAllMean$T_C.SubstRate.heavy/mutSpecAllMean$subsratesum
mutSpecAllMean$T_G.SubstRate.Percent.heavy=mutSpecAllMean$T_G.SubstRate.heavy/mutSpecAllMean$subsratesum
```

### Delete bad rows and take fishes that have codons and temperature

```{r}
mutSpecAllMean = mutSpecAllMean[!is.na(mutSpecAllMean$temperature),]
```

### take 10% of cold and warm fishes 

```{r}
quant = quantile(mutSpecAllMean$temperature, probs = c(0.1, 0.9))

cold = mutSpecAllMean[mutSpecAllMean$temperature <= quant[[1]],]; nrow(cold)

warm = mutSpecAllMean[mutSpecAllMean$temperature >= quant[[2]],]; nrow(warm)


cold_fish = apply(as.matrix(cold[,22:33]), 2, mean)
warm_fish = apply(as.matrix(warm[,22:33]), 2, mean)
```

### Change to better names

```{r}
names(cold_fish) = c('A_C','A_G','A_T','C_A','C_G','C_T','G_A','G_C','G_T','T_A','T_C','T_G')
names(warm_fish) = c('A_C','A_G','A_T','C_A','C_G','C_T','G_A','G_C','G_T','T_A','T_C','T_G')
```

## Simmulation

### INITIALIZE PARAMETERS:

```{r}
Final = data.frame()
GenomeLength = 10000  
SimulationLengthNumberOfGenerations = 600000
```

### Body of simmulation

```{r}
for (MutSpecProb in c('cold-water fish','warm-water fish'))
{
  ### choose initial nucleotide frequencies: equal to 25% if InitGenome == 1 or random if InitGenome > 1
  for (InitGenome in 1:10)
  {  #InitGenome = 2
    if (InitGenome == 1) {frA = frG = frC = frT = 0.25} 
    if (InitGenome >  1) 
    {
      frA = runif(1); frG = runif(1);
      frC = runif(1); frT = runif(1);  
      Summa = frA+frT+frG+frC
      frA = frA/Summa; frG = frG/Summa; frC = frC/Summa; frT = frT/Summa; 
      frA+frG+frC+frT # should be 1
    }
    
    ### make a genome
    genome = sample(c(rep('A',round(frA*GenomeLength)),rep('T',round(frT*GenomeLength)),rep('G',round(frG*GenomeLength)),rep('C',round(frC*GenomeLength))))
    length(genome) # should be equal GenomeLength
    
    ##### B: DEFINE MUTATIONAL SPECTRUM
    # Take MutSpec from the two values: cold_fish and worm_fish
    
    if (MutSpecProb == 'cold-water fish')
    { #cold_fish
      VecMutSpec = c(  
      "T","A",cold_fish[['T_A']],
      "T","C",cold_fish[['T_C']], 
      "T","G",cold_fish[['T_G']],
      "A","T",cold_fish[['A_T']],
      "A","C",cold_fish[['A_C']],
      "A","G",cold_fish[['A_G']], 
      "C","T",cold_fish[['C_T']],
      "C","A",cold_fish[['C_A']],
      "C","G",cold_fish[['C_G']],
      "G","T",cold_fish[['G_T']],
      "G","A",cold_fish[['G_A']],
      "G","C",cold_fish[['G_C']])} 
    
    if (MutSpecProb == 'warm-water fish')
    { #warm_fish
      VecMutSpec = c(  
        "T","A",warm_fish[['T_A']],
        "T","C",warm_fish[['T_C']], 
        "T","G",warm_fish[['T_G']],
        "A","T",warm_fish[['A_T']],
        "A","C",warm_fish[['A_C']],
        "A","G",warm_fish[['A_G']], 
        "C","T",warm_fish[['C_T']],
        "C","A",warm_fish[['C_A']],
        "C","G",warm_fish[['C_G']],
        "G","T",warm_fish[['G_T']],
        "G","A",warm_fish[['G_A']],
        "G","C",warm_fish[['G_C']])} 
    
    MutSpec = data.frame(matrix(VecMutSpec, ncol = 3, nrow = 12, byrow = TRUE))
    names(MutSpec) = c('From','To','Prob')
    MutSpec$Prob = as.numeric(as.character(MutSpec$Prob))
    sum(MutSpec$Prob)
    
    ExpectedFrA = sum(MutSpec[MutSpec$To == 'A',]$Prob)/sum(MutSpec[MutSpec$From == 'A',]$Prob)
    ExpectedFrT = sum(MutSpec[MutSpec$To == 'T',]$Prob)/sum(MutSpec[MutSpec$From == 'T',]$Prob)
    ExpectedFrG = sum(MutSpec[MutSpec$To == 'G',]$Prob)/sum(MutSpec[MutSpec$From == 'G',]$Prob)
    ExpectedFrC = sum(MutSpec[MutSpec$To == 'C',]$Prob)/sum(MutSpec[MutSpec$From == 'C',]$Prob)
    Summa = ExpectedFrA + ExpectedFrT + ExpectedFrG + ExpectedFrC
    ExpectedFrA = ExpectedFrA/Summa
    ExpectedFrT = ExpectedFrT/Summa 
    ExpectedFrG = ExpectedFrG/Summa
    ExpectedFrC = ExpectedFrC/Summa

    
    sum(MutSpec$Prob) 
    
    for (i in 1:nrow(MutSpec))
    { # i = 1
      if (i == 1) {MutSpec$RulletFrom[i] = 0}
      MutSpec$RulletTo[i] = sum(MutSpec[seq(1:i),]$Prob)
      if (i  > 1) {MutSpec$RulletFrom[i] = MutSpec$RulletTo[i-1]}
    }
    
    ##### C: MUTATE AND SAVE NUCLEOTIDE CONTENT EVERY 1000 GENERATIONS
    for (gener in 1:SimulationLengthNumberOfGenerations)
    {
      ### 3: choose a random position in a genome
      RandomPos = sample(1:length(genome), 1)  
      NucInRandomPos = genome[RandomPos];
      
      ### 2: choose a random mutation
      Rullet = runif(1)
      PotentialSubstitution = MutSpec[MutSpec$RulletFrom <= Rullet & MutSpec$RulletTo > Rullet,]
      PotentialSubstitution$To = as.character(PotentialSubstitution$To)
      ### 3: mutation happens
      if (nrow(PotentialSubstitution) == 1)
      {
        if (NucInRandomPos == PotentialSubstitution$From)  {genome[RandomPos] = PotentialSubstitution$To}
      }
      
      ### 4: print out every 1000 generations
      if ((gener %% 1000) == 0)
      {
        Res = data.frame(table(genome))
        Res = data.frame(t(Res[order(Res$genome),]))
        Res = Res[2,]
        Res$Gener = gener
        Res$InitGenome = InitGenome
        Res$MutSpecProb = MutSpecProb
        names(Res)= c('A','C','G','T','Gener','InitGenome','MutSpecProb')
        Final = rbind(Final,Res)
      }
    }}}
```

### Count FrX

```{r}
row.names(Final) = 1:nrow(Final)
Final$FrA = as.numeric(as.character(Final$A))/length(genome)
Final$FrT = as.numeric(as.character(Final$T))/length(genome)
Final$FrG = as.numeric(as.character(Final$G))/length(genome)
Final$FrC = as.numeric(as.character(Final$C))/length(genome)
```

## Block to draw simmulation pictures

```{r}
ColorA = rgb(1,0,0,0.3)
ColorT = rgb(0,1,0,0.3)
ColorG = rgb(0,0,1,0.3)
ColorC = rgb(0,1,1,0.3)


Final = Final[Final$Gener <= 600000,]
###: PLOT

pdf("../Figures/SimulationFishes.pdf",  width = 20, height = 10) # dev.off()
par(mfrow=c(2,2))

for (InitGenome in 1:10)
{
  plot(Final[Final$InitGenome == InitGenome & Final$MutSpecProb == 'cold-water fish',]$Gener,Final[Final$InitGenome == InitGenome & Final$MutSpecProb == 'cold-water fish',]$FrA, ylim = c(0,0.6), pch = '.', col = ColorA, main = '', xlab = '', ylab = ''); par(new=TRUE)
  plot(Final[Final$InitGenome == InitGenome & Final$MutSpecProb == 'cold-water fish',]$Gener,Final[Final$InitGenome == InitGenome & Final$MutSpecProb == 'cold-water fish',]$FrT, ylim = c(0,0.6), pch = '.', col = ColorT, main = '', xlab = '', ylab = ''); par(new=TRUE)
  plot(Final[Final$InitGenome == InitGenome & Final$MutSpecProb == 'cold-water fish',]$Gener,Final[Final$InitGenome == InitGenome & Final$MutSpecProb == 'cold-water fish',]$FrG, ylim = c(0,0.6), pch = '.', col = ColorG, main = '', xlab = '', ylab = ''); par(new=TRUE)
  plot(Final[Final$InitGenome == InitGenome & Final$MutSpecProb == 'cold-water fish',]$Gener,Final[Final$InitGenome == InitGenome & Final$MutSpecProb == 'cold-water fish',]$FrC, ylim = c(0,0.6), pch = '.', col = ColorC, main = 'Cold-water, neutral equilibrium, heavy chain', xlab = '', ylab = ''); par(new=TRUE)
}

A =  0.1843125066 ; G = 0.2036486321 
T = 0.5415029364; C=0.07053592541 # data from Valerian

abline(h = A, col = ColorA, lwd = 3, lt = 2)
abline(h = T, col = ColorT, lwd = 3, lt = 2)
abline(h = G, col = ColorG, lwd = 3, lt = 2)
abline(h = C, col = ColorC, lwd = 3, lt = 2)

legend(1,0.6, legend = c('A','T','G','C'), col = c(ColorA, ColorT, ColorG, ColorC), pch = 16)

for (InitGenome in 1:10)
{
  plot(Final[Final$InitGenome == InitGenome & Final$MutSpecProb == 'warm-water fish',]$Gener,Final[Final$InitGenome == InitGenome & Final$MutSpecProb == 'warm-water fish',]$FrA, ylim = c(0,0.6), pch = '.', col = ColorA, main = '', xlab = '', ylab = ''); par(new=TRUE)
  plot(Final[Final$InitGenome == InitGenome & Final$MutSpecProb == 'warm-water fish',]$Gener,Final[Final$InitGenome == InitGenome & Final$MutSpecProb == 'warm-water fish',]$FrT, ylim = c(0,0.6), pch = '.', col = ColorT, main = '', xlab = '', ylab = ''); par(new=TRUE)
  plot(Final[Final$InitGenome == InitGenome & Final$MutSpecProb == 'warm-water fish',]$Gener,Final[Final$InitGenome == InitGenome & Final$MutSpecProb == 'warm-water fish',]$FrG, ylim = c(0,0.6), pch = '.', col = ColorG, main = '', xlab = '', ylab = ''); par(new=TRUE)
  plot(Final[Final$InitGenome == InitGenome & Final$MutSpecProb == 'warm-water fish',]$Gener,Final[Final$InitGenome == InitGenome & Final$MutSpecProb == 'warm-water fish',]$FrC, ylim = c(0,0.6), pch = '.', col = ColorC, main = 'Warm-water, neutral equilibrium, heavy chain', xlab = '', ylab = ''); par(new=TRUE)
}

A = 0.1570029522; G=0.2119039825; 
T = 0.5186136993; C = 0.1124793649 # data from Valerian:

abline(h = A, col = ColorA, lwd = 3, lt = 2)
abline(h = T, col = ColorT, lwd = 3, lt = 2)
abline(h = G, col = ColorG, lwd = 3, lt = 2)
abline(h = C, col = ColorC, lwd = 3, lt = 2)
legend(1,0.6, legend = c('A','T','G','C'), col = c(ColorA, ColorT, ColorG, ColorC), pch = 16)

dev.off()
```

## ExpectedbvsOberved analysys

### take data from Valerian's diffur

```{r}
expected_cold = data.frame(0.1843125066, 0.2036486321, 0.5415029364, 0.07053592541)
names(expected_cold) =c('FrA','FrG','FrT','FrC')

expected_warm = data.frame(0.1570029522, 0.2119039825, 0.5186136993, 0.1124793649)
names(expected_warm) =c('FrA','FrG','FrT','FrC')
```

### reverse to put in graph

```{r}
expected_all = cbind(expected_cold, expected_warm)
expected_all = t(as.matrix(expected_all))
```

### Write types and colnames

```{r}
expected_all = cbind(expected_all,c('cold_fish','cold_fish','cold_fish','cold_fish','warm_fish','warm_fish','warm_fish','warm_fish'))
expected_all = cbind(expected_all, c('FrA','FrG','FrT','FrC','FrA','FrG','FrT','FrC'))
colnames(expected_all) = c('expected', 'type_of_fish','mutation')
```

### Observed mutations df

```{r}
cold_fish = (apply(as.matrix(cold[,5:8]), 2, mean))
warm_fish = (apply(as.matrix(warm[,5:8]), 2, mean))

observed = cbind(t(as.matrix(cold_fish)), t(as.matrix(warm_fish)))
observed = t(as.matrix(observed))
observed = cbind(observed, c('FrA', 'FrT', 'FrG', 'FrC', 'FrA', 'FrT', 'FrG', 'FrC')) 
observed = cbind(observed, c('cold_fish','cold_fish','cold_fish','cold_fish','warm_fish','warm_fish','warm_fish','warm_fish'))
colnames(observed) = c('observed','mutation','type_of_fish')
```

### Merging to draw

```{r}
expectedvsoberved = merge(expected_all, observed, by = c('mutation','type_of_fish'))
expectedvsoberved$observed =as.numeric(as.character(expectedvsoberved$observed))
expectedvsoberved$expected =as.numeric(as.character(expectedvsoberved$expected))
```

### Draw figure 3
```{r}
pdf("../Figures/ExpectedvsObserved.pdf")

exvsobs = ggplot(data = expectedvsoberved, aes(x = observed, y = expected, group=type_of_fish, col = type_of_fish))+
  geom_point(size = 3.5)+
  geom_abline(col = 'gray3', linetype="longdash", size = 0.6)+
  theme_bw()+
  scale_color_manual(name="Type of Fish", labels = c('Cold fish', 'Warm fish'), values = c('cold_fish'='deepskyblue4', 'warm_fish' = 'firebrick3'))+
  geom_line(aes(group = mutation), col = 'black', size = 0.7)+
  geom_text(aes(label=mutation),hjust=-0.40, vjust=-0.40)+
  xlim(0, 0.55)+
  ylim(0,0.55)+
  labs(x = 'Observed Nucleotide Content',y = 'Expected Nucleotide Content')
  

print(exvsobs)

dev.off()
```

